Project 1

### 2018-09-10 Project1/report/ms.tex
Updates from ShareLaTeX

\section{Introduction}
\label{sec:introduction}
This study on generalized and tailored numerical algorithms will focus on how a differential equation is solved numerically most efficiently while not reducing the accuracy of the calculations. By comparing the run time and observing the numerical error of the calculations . To specify, the equation in question is the one dimensional Poisson equation:
%
\begin{equation*}
    {\nabla^{2}} \Phi = -4\pi \rho(r),
\end{equation*}
%
where $\Phi$ is the electrostatic potential generated by a localized charge distribution and $\rho$ is the charge density. This equation is used in electromagnetism, but in this case it was only necessary to analyze a normalized equation on the same form. Our normalized one dimensional Poisson equation can then be written as:
%
\begin{equation*}\label{eq:-u''}
    -\frac{du^2}{dx^2} = f(x).
\end{equation*}

@@

\section{Method}
\label{sec:method}
For the study we consider a solution for the source term $f(x)$ to be
%
\begin{gather}\label{eq:f(x)}
    f(x) = 100e^{-10x},
\end{gather}
%
which has a closed form solution 
%
\begin{gather}\label{eq:u(x)}
u(x) = 1-(1-e^{10})x-e^{-10x},
\end{gather}
%
which yields $-f(x)$ when taking the second derivative.

We consider the one dimensional Poisson equation to uphold Dirichlet boundary conditions; $u(0)=u(1)=0$, so $x \in [0,1]$. The second order derivative $u(x)$ can be represented as
%
\begin{gather*}
 -\frac{du^2}{dx^2} = -\left(\frac{u(x+h) + u(x-h) - 2u(x)}{h^2} + \mathcal{O}(h^2)\right),
\end{gather*}
%
where h is the step size. By approximating the above equation and discretizing over x we get
%
\begin{gather*}
    -\frac{du^2}{dx^2} = -\frac{u_{i+1} + u_{i-1} - 2u_i}{h^2} = f(x_i) = f_i.
\end{gather*}
%
To simplify the calculations and reduce the FLOPS we multiply the above equation by $h^2$ and rewrite $f'_i = h^2f_i = h^2 100 e^{-10x_i}$, such that the final discretized equation for solving equation \eqref{eq:-u''} is
%
\begin{gather}\label{eq:u_discretized}
    u_{i+1} + u_{i-1} - 2u_i = f'_i.
\end{gather}
%
The left hand side of the above equation can be represented as a matrix vector multiplication.

\[ \boldsymbol{Au} =
\begin{array}{c}
\begin{bmatrix}\label{eq:Au=f}
b_1     & c_1           & 0         & \dots     & \dots     & 0 \\
a_1     & b_2           & c_2       & 0         & \dots     & \dots \\
0       & a_2           & b_3       & c_3       & 0     & \dots\\
\dots  &  0            & a_3       & b_4       & \dots    & 0 \\
\dots  & \dots        & 0 & \dots    & \dots    & c_{n-1} \\
0       & \dots         & \dots         & 0         & a_{n-1}   & b_n   \\
\end{bmatrix}
\begin{bmatrix}
u_1 \\
u_2 \\
\dots \\
\dots \\
\dots \\
u_n
\end{bmatrix}
=
\begin{bmatrix}
f'_1 \\
f_2 \\
\dots \\
\dots \\
\dots \\
f'_n
\end{bmatrix}
\end{array}
= \boldsymbol{f'_i}
\].

%\begin{bmatrix}
%\Tilde{f}_1 \\
%\Tilde{f}_2 \\
%\dots \\
%\dots \\
%\dots \\
%\Tilde{f}_n
%\end{bmatrix}
%\end{array}

\subsection{Generalized tridiagonal matrix}

@@

>>>>

\section{Introduction}
\label{sec:introduction}
This study on generalized and tailored numerical algorithms will focus on how a differential equation is solved numerically most efficiently while not reducing the accuracy of the calculations. By comparing the run time and observing the numerical error of the calculations . To specify, the equation in question is the one dimensional Poisson equation:
%
\begin{equation*}
    {\nabla^{2}} \Phi = -4\pi \rho(r),
\end{equation*}
%
where $\Phi$ is the electrostatic potential generated by a localized charge distribution and $\rho$ is the charge density. This equation is used in electromagnetism, but in this case it was only necessary to analyze a normalized equation on the same form. Our normalized one dimensional Poisson equation can then be written as:
%
\begin{equation}\label{eq:-u''}
    -\frac{du^2}{dx^2} = f(x).
\end{equation}

@@

\section{Method}
\label{sec:method}
For the study we consider a solution for the source term $f(x)$ to be
%
\begin{gather*}\label{eq:f(x)}
    f(x) = 100e^{-10x},
\end{gather*}
%
which has a closed form solution 
%
\begin{gather}\label{eq:u(x)}
u(x) = 1-(1-e^{10})x-e^{-10x},
\end{gather}
%
which yields $-f(x)$ when taking the second derivative.

We consider the one dimensional Poisson equation to uphold Dirichlet boundary conditions; $u(0)=u(1)=0$, so $x \in [0,1]$. The second order derivative $u(x)$ can be represented as
%
\begin{gather}\label{eq:du/dx}
 -\frac{du^2}{dx^2} = -\left(\frac{u(x+h) + u(x-h) - 2u(x)}{h^2} + \mathcal{O}(h^2)\right),
\end{gather}
%
where h is the step size.

\subsection{Numerical representation}
By approximating equation \eqref{eq:du/dx} and discretizing over x we get
%
\begin{gather*}
    -\frac{du^2}{dx^2} = -\frac{u_{i+1} + u_{i-1} - 2u_i}{h^2} = f(x_i) = f_i.
\end{gather*}
%
To simplify the calculations and reduce the FLOPS, we multiply the above equation by $h^2$ and rewrite $f'_i = h^2f_i = h^2 100 e^{-10x_i}$, such that the final discretized equation for solving equation \eqref{eq:-u''} is
%
\begin{gather}\label{eq:u_discretized}
    -\left(u_{i-1} - 2u_i + u_{i+1}\right) = f'_i.
\end{gather}
%
The linear combination above can be represented by a matrix vector multiplication. The matrix will consist of the coefficients of the $u_i$'s along the central, upper and lower diagonal, making up a tridiagonal matrix. The coefficient, after taking the outer sign into  consideration, are -1, 2 and -1 respectively. The vector will consist of the values of the $u_i$'s. The matrix representation will therefore be on the following form:
%
\[ \boldsymbol{Au} =
\begin{array}{c}
\begin{bmatrix}\label{eq:Au=f}
b_1     & c_1           & 0         & \dots     & \dots     & 0 \\
a_1     & b_2           & c_2       & 0         & \dots     & \dots \\
0       & a_2           & b_3       & c_3       & 0     & \dots\\
\dots  &  0            & a_3       & b_4       & \dots    & 0 \\
\dots  & \dots        & 0 & \dots    & \dots    & c_{n-1} \\
0       & \dots         & \dots         & 0         & a_{n-1}   & b_n   \\
\end{bmatrix}
\begin{bmatrix}
u_1 \\
u_2 \\
\dots \\
\dots \\
\dots \\
u_n
\end{bmatrix}
=
\begin{bmatrix}
f'_1 \\
f_2 \\
\dots \\
\dots \\
\dots \\
f'_n
\end{bmatrix}
\end{array}
= \boldsymbol{f'_i}.
\]
%
Since we are considering fixed end points for $u$ we do not include $i = 0$ and $i = n+1$ in the calculations for updating the $u_i$'s.
%\begin{bmatrix}
%\Tilde{f}_1 \\
%\Tilde{f}_2 \\
%\dots \\
%\dots \\
%\dots \\
%\Tilde{f}_n
%\end{bmatrix}
%\end{array}

@@

\begin{figure}[H]
    \centering
    \includegraphics[size=0.25\textwidth]{plot1d.png}
    \caption{}
    \label{fig1:1d}
\end{figure}
